Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-21 appendix-EPQ.tex
modify

We generate register automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}. Here $C_1 = \{ \textit{call}(\textit{put},a,\top)$, $\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},b),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2$ $= C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\top) \}$.

@@

We generate register automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}. Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{call}(\textit{rm},b) + \textit{ret}(\textit{rm},b) \}$.

@@

We generate register automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}. Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,\top) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_4 = C_3 \cup \{ \textit{call}(\textit{rm},b) \}$, $C_5 = C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_6 = C_5 \cup \{ \textit{call}(\textit{rm},b) \}$.

@@

We generate register automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}. Here $C_1 = C \cup \{ \textit{call}(\textit{rm},b) \}$, and $C_2 = C \cup \{ \textit{ret}(\textit{put},b,\top), \textit{call}(\textit{rm},a), \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},d,\top),\textit{ret}(\textit{put},d,\top), \textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.

@@

@@

\begin{restatable}{lemma}{UVSetHasMatchedPutandRm}
\label{lemma:UVSet has matched put and rm}
Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$. Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. Assume that $G$ has no cycle going through $x$. Then, $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.
\end{restatable}
\begin {proof}

@@

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] For each $1 \leq i \leq j$, method events of $d_i$ belongs to $\textit{UVSet}_i(e,x)$.

@@

Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction. Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap. Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{UVSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$. Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$. Since method events of $d_{\textit{i+1}}$ is in $\textit{UVSet}_{\textit{i+1}}(e,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$. Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that for each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.

@@

Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$. Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. Assume that $G$ has no cycle going through $x$. Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}(e,x)$.
\end{restatable}

@@

\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_1(e,x)$ is an operation of item $d$. %(according to the definition of $\textit{UVSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).

We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ method event or a $\textit{rm}$ method event. $t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_1(e,x)$: $o \in \textit{UVSet}_1(e,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.

@@

\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x) )$. By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,\textit{pri}_x) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.

@@

\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x) )$. We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.

@@

\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.

@@

\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$. Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_{\textit{j+1}}(e,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$ is an operation of item $d_{\textit{j+1}}$. We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ method event or a $\textit{rm}$ method event. $t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$: $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{UVSet}_j(e,x)$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

@@

With Lemma \ref{lemma:UVSet has matched put and rm} and Lemma \ref{lemma:Rmx does not happen before UVSet for EPQ1Lar}, we can now prove Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}.

\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
\label{lemma:Lin Equals Constraint for EPQ1Lar}
Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$. Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$. $e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$, if and only if $G$ has no cycle going through $x$.
\end{restatable}

@@

To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$. Let $u$, $v$ and $w$ be the sequences of method events in $\seqPQ_1^{>}$, and let $U$, $V$ and $W$ be the set of method events of $u$, $v$ and $w$, respectively. Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$. It is obvious that the priority of each $d_i$ is smaller than $\textit{pri}_x$. Then our proof proceeds as follows:

@@

By Lemma \ref{lemma:UVSet has matched put and rm}, we can see that $E_u \cup E_v$ contains matched $\textit{put}$ and $\textit{rm}$ operations. It remains to prove that for $E_u$, $\{ \textit{put}(x,\textit{pri}_x) \}$, $E_v$, $\{ \textit{rm}(x) \}$, $E_w$, no elements of the latter set happens before elements of the former set. We prove this by showing that all the following cases are impossible:

@@

Let us begin to represent register automata that is used for capture the existence of a data-differentiated execution $e$, $e$ has a $\_$-projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$. By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$. There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$. For each of them, we generate a register automaton.

For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}. Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{call}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{call}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

@@

Let $o$ be the right-most gap-point of $x$. We locate linearization points of each method event as follows:

@@

\item[-] Locate linearization points of method event of each $y \in \textit{Items}(e,\textit{pri})$ (except for $x$) just after the call action of the method event.

@@

According to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar} (Here we temporarily forget the existence of $y$), the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $x$. Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched method event before $\textit{rm}(x)$ with smaller priority. Since the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},x)$.

Therefore, for every sequence $l = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched method events of priority smaller than $\textit{pri}$, or $w$ contains $\textit{put}(y,\textit{pri})$ or $\textit{rm}(y)$. In both cases, $l \notin \textit{MS}(\seqPQ_1^{=})$.

@@

There are many enumerations of method events of $a$, $b$ and $a_1$ that may makes $a <_{\textit{pb}}^* b$. The following lemma states that with the help of gap-points, the number of potential enumerations can be further reduced into only five.

\begin{restatable}{lemma}{FiveEnmuerationisEnoughForEPQOneEqual}
\label{lemma:five enumeration is enough for EPQ1Equal}
Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$. Let $a$ and $b$ be items with maximal priority $\textit{pri}$. Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\textit{pri})$ or $\textit{call}(\textit{rm},a)$. Then, there are five possible enumeration of method events of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
\end{restatable}
\begin {proof}

@@

\begin {proof}
Since $\textit{last}(e) = \seqPQ_2^{>}$, the actions with maximal priority in $e$ is some unmatched $\textit{put}$. Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of method events that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the guard of $\textit{MS}(\seqPQ_2^{>})$. This completes the proof of this lemma. \qed
\end {proof}

@@

\item[-] For other method events, locate their linearization points at an arbitrary location after its call action and before its return action.
\end{itemize}

@@

To prove the $\textit{if}$ direction, assume that there is such a cycle. Assume by contradiction that $e \sqsubseteq \textit{MS}(\seqPQ_3)$, and let $U$ and $V$ be the set of operations in $u$ and $v$. Let the cycle be $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow o \rightarrow d_1$ in $G$. Since $d_m \rightarrow o$, $\textit{put}(d_m,\_)$ happens before $o$, and it is easy to see that $\textit{put}(d_m,\_)$ is in $U$. Since $U$ contains matched $\textit{put}$ and $\textit{rm}$, we can see that operations of $d_m$ is in $U$. Similarly, we can see that method events of $d_{\textit{m-1}},\ldots,d_1$ is in $U$. If $\textit{rm}(d_1)$ does not exists, then this contradicts that $U$ contains matched $\textit{put}$ and $\textit{rm}$. Else, if $\textit{rm}(d_1)$ exists, since $o$ happens before $\textit{rm}(d_1)$, we can see that $\textit{rm}(d_1) \in V$, which contradicts that $\textit{rm}(d_1) \in U$. This completes the proof of the $\textit{if}$ direction.

@@

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Given a sequential execution $l$ of multi-set, $\textit{noDE}(l)$ is satisfied when each method event of $l$ is not $\textit{delete}(\textit{empty})$.

@@

\item[-] Given a sequential execution $l$ of multi-set, $l \in \textit{Insert}^*$ is satisfied when each method event of $l$ is a $\textit{insert}$ event.
\end{itemize}

>>>>

We generate register automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}. Here $C_1 = \{ \textit{call}(\textit{put},a,\textit{true})$, $\textit{ret}(\textit{put},a,\textit{true}), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},b),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2$ $= C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$.

@@

We generate register automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}. Here $C_1 = \{ \textit{call}(\textit{put},a,\textit{true}),\textit{ret}(\textit{put},a,\textit{true}), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{call}(\textit{rm},b) + \textit{ret}(\textit{rm},b) \}$.

@@

We generate register automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}. Here $C_1 = \{ \textit{call}(\textit{put},a,\textit{true}),\textit{ret}(\textit{put},a,\textit{true}), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_4 = C_3 \cup \{ \textit{call}(\textit{rm},b) \}$, $C_5 = C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_6 = C_5 \cup \{ \textit{call}(\textit{rm},b) \}$.

@@

We generate register automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}. Here $C_1 = C \cup \{ \textit{call}(\textit{rm},b) \}$, and $C_2 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{call}(\textit{rm},a), \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},d,\textit{true}),\textit{ret}(\textit{put},d,\textit{true}), \textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.

@@

Therefore, from now on, it is safe to consider only the data-differentiated sequences with only one maximal priority.

@@

\begin{restatable}{lemma}{UVSetHasMatchedPutandRm}
\label{lemma:UVSet has matched put and rm}
Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$. Let $G$ be the graph representing the left-right constraint of $x$. Assume that $G$ has no cycle going through $x$. Then, $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.
\end{restatable}
\begin {proof}

@@

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] For each $1 \leq i \leq j$, operations of $d_i$ belongs to $\textit{UVSet}_i(e,x)$.

@@

Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction. Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap. Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{UVSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$. Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$. Since operations of $d_{\textit{i+1}}$ is in $\textit{UVSet}_{\textit{i+1}}(e,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$. Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that for each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.

@@

Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$. Let $G$ be the graph representing the left-right constraint of x. Assume that $G$ has no cycle going through $x$. Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}(e,x)$.
\end{restatable}

@@

\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_1(e,x)$ is an operation of item $d$.

We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation. $t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_1(e,x)$: $o \in \textit{UVSet}_1(e,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,p)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,p)$.

@@

\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,p) )$. By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,p) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.

@@

\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,p) )$. We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.

@@

\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,p)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,p)$.

@@

\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$. Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_{\textit{j+1}}(e,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$ is an operation of item $d_{\textit{j+1}}$. We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation. $t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$: $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{UVSet}_j(e,x)$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

@@

With Lemma \ref{lemma:UVSet has matched put and rm} and Lemma \ref{lemma:Rmx does not happen before UVSet for EPQ1Lar}, we can now prove Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}, which give the equivalent characterization of non $\mathsf{MatchedMaxPriority}$-linearizable executions.

\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
\label{lemma:Lin Equals Constraint for EPQ1Lar}
Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$. Let $G$ be the graph representing the left-right constraint of $x$. $e$ is $\mathsf{MatchedMaxPriority}$-linearizable, if and only if $G$ has no cycle going through $x$. 
\end{restatable}

@@

To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq l$ and $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds. Then $l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, and let $U$, $V$ and $W$ be the set of operations of $u$, $v$ and $w$, respectively. Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$. It is obvious that the priority of each $d_i$ is smaller than $p$. Then our proof proceeds as follows:

@@

By Lemma \ref{lemma:UVSet has matched put and rm}, we can see that $E_u \cup E_v$ contains matched $\textit{put}$ and $\textit{rm}$ operations. It remains to prove that for $E_u$, $\{ \textit{put}(x,p) \}$, $E_v$, $\{ \textit{rm}(x) \}$, $E_w$, no elements of the latter set happens before elements of the former set. We prove this by showing that all the following cases are impossible:

@@

Let us begin to represent register automata that is used for capture the existence of a data-differentiated execution $e$, let $p$ be a maximal priority of $e$, and $e'$ be the projection of $e$ into values with priorities comparable with $p$. We also require that $\mathsf{Has\text{-}MatchedMaxPriority}(e')$ holds and there exists a cycle going through the value with maximal priority in $e'$. By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$. There are four possible enumeration of call and return actions of $\textit{put}(b,p)$ and $\textit{rm}(b)$. For each of them, we generate a register automaton.

For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put},b,p) \cdot \textit{call}(\textit{rm},b) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}. Here $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,<r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

@@

Let $o$ be the right-most gap-point of $x$. We locate linearization points of each operation as follows:

@@

\item[-] Locate linearization points of operation of each $y \in \textit{Items}(e,\textit{pri})$ (except for $x$) just after the call action of the operation.

@@

According to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar} (Here we temporarily forget the existence of $y$), the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $x$. Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched operation before $\textit{rm}(x)$ with smaller priority. Since the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},x)$.

Therefore, for every sequence $l = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority smaller than $\textit{pri}$, or $w$ contains $\textit{put}(y,\textit{pri})$ or $\textit{rm}(y)$. In both cases, $l \notin \textit{MS}(\seqPQ_1^{=})$.

@@

There are many enumerations of operations of $a$, $b$ and $a_1$ that may makes $a <_{\textit{pb}}^* b$. The following lemma states that with the help of gap-points, the number of potential enumerations can be further reduced into only five.

\begin{restatable}{lemma}{FiveEnmuerationisEnoughForEPQOneEqual}
\label{lemma:five enumeration is enough for EPQ1Equal}
Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$. Let $a$ and $b$ be items with maximal priority $\textit{pri}$. Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\textit{pri})$ or $\textit{call}(\textit{rm},a)$. Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
\end{restatable}
\begin {proof}

@@

\begin {proof}
Since $\textit{last}(e) = \seqPQ_2^{>}$, the actions with maximal priority in $e$ is some unmatched $\textit{put}$. Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of operations that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the guard of $\textit{MS}(\seqPQ_2^{>})$. This completes the proof of this lemma. \qed
\end {proof}

@@

\item[-] For other operations, locate their linearization points at an arbitrary location after its call action and before its return action.
\end{itemize}

@@

To prove the $\textit{if}$ direction, assume that there is such a cycle. Assume by contradiction that $e \sqsubseteq \textit{MS}(\seqPQ_3)$, and let $U$ and $V$ be the set of operations in $u$ and $v$. Let the cycle be $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow o \rightarrow d_1$ in $G$. Since $d_m \rightarrow o$, $\textit{put}(d_m,\_)$ happens before $o$, and it is easy to see that $\textit{put}(d_m,\_)$ is in $U$. Since $U$ contains matched $\textit{put}$ and $\textit{rm}$, we can see that operations of $d_m$ is in $U$. Similarly, we can see that operations of $d_{\textit{m-1}},\ldots,d_1$ is in $U$. If $\textit{rm}(d_1)$ does not exists, then this contradicts that $U$ contains matched $\textit{put}$ and $\textit{rm}$. Else, if $\textit{rm}(d_1)$ exists, since $o$ happens before $\textit{rm}(d_1)$, we can see that $\textit{rm}(d_1) \in V$, which contradicts that $\textit{rm}(d_1) \in U$. This completes the proof of the $\textit{if}$ direction.

@@

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Given a sequential execution $l$ of multi-set, $\textit{noDE}(l)$ is satisfied when each operation of $l$ is not $\textit{delete}(\textit{empty})$.

@@

\item[-] Given a sequential execution $l$ of multi-set, $l \in \textit{Insert}^*$ is satisfied when each operation of $l$ is a $\textit{insert}$ event.
\end{itemize}
